Author:  Benjamin Müller
Course:   Data Vizualization 
Exercise: Project
Date of talk: 11.12.2019


#Settings in advance
Please change the working directory to the working directory you're working on.

```{r}
setwd("C:/Users/deyno/Documents/R Workspace/DataVizProjekt - DatenKlima")
set.seed(666) #probably not necessary
rm(list = ls())
options(scipen = 999) #disabling scientific e+10 notification
```


#install.packages

```{r}
install.packages("ggplot2")
install.packages("eurostat")
install.packages("reshape2")
install.packages("ggrepel")
install.packages("dplyr")
```

#load.packages

```{r}
library("ggplot2")
library("eurostat")
library("reshape2")
library("ggrepel")
library("dplyr")
library(tidyverse)
library(HSAUR3)
library(factoextra)
```


#--load and adjust data sets--

Datasets are loaded from the website, because the automated tool get_eurostat from eurostat package shortens the labels and it's easier to choose variables on the website. Datasets are loaded from csv.

```{r}
env_ac_ainah_r2 <- read.csv("./Data/Air emissions accounts by NACE Rev. 2 activity (env_ac_ainah_r2)/env_ac_ainah_r2_1_Data.csv")
```

Have a look at the data set and change data type, remove the NAs etc.


```{r}
str(env_ac_ainah_r2)

env_ac_ainah_r2$Value <- levels(env_ac_ainah_r2$Value)[env_ac_ainah_r2$Value]

env_ac_ainah_r2$Value <- gsub(",","",env_ac_ainah_r2$Value) 
     # 1000 are seperated by comma, e.g. 1,000,000 for a million
     # This has to be deleted for the as.numeric to work properly

env_ac_ainah_r2$Value <- as.double(env_ac_ainah_r2$Value) #NAs introduced by coercion!

sum(is.na(env_ac_ainah_r2$Value))
#There are 198 NAs here

#Simple imputation method: Replace NAs by 0

env_ac_ainah_r2$Value <- replace(env_ac_ainah_r2$Value, is.na(env_ac_ainah_r2$Value), 0)

#Check the sum of NAs and the resulting sum is zero.


#shorten the labels of the factors---

#look at the factor names of the account of greenhous gases
levels(env_ac_ainah_r2$NACE_R2) 

#too long descriptions shall be shortened, hopefully without changing the meaning
#Careful approch by now
levels(env_ac_ainah_r2$NACE_R2) [levels(env_ac_ainah_r2$NACE_R2) == "Activities of households as employers; undifferentiated goods- and services-producing activities of households for own use"] <- "Activities of households es employers"

#Look at the names of the states

levels(env_ac_ainah_r2$GEO) [levels(env_ac_ainah_r2$GEO) == "Germany (until 1990 former territory of the FRG)"] <- "Germany"
```

# Question: What are the most important drivers of greenhouse gase emission in the European Union?

To answer this question the year 2017 is taken as an example and simply the drivers or the accounts
of greenhouse emission are shown. So the result will be just one dimension.

A visualization, which might be suitable is the barplot. For better readibiity the barplot shall be
ordered.

First the data has to be subsetted, since we just want to have all countries and just the year 2017.

```{r}
env_ac_ainah_r2.total.2017 <- subset(env_ac_ainah_r2, (TIME == 2017) & (GEO == "European Union - 28 countries") & !(NACE_R2 == "Total - all NACE activities"))
```

#Make a barplot

```{r}
plot.2017.total.account.barplot <- ggplot(env_ac_ainah_r2.total.2017, aes(x=reorder(NACE_R2,Value), y=Value, fill=NACE_R2))+
  geom_bar(stat = "identity", color = "brown") +
  coord_flip() +
  theme(legend.position = "none") +
  xlab("Account of Greenhous Gases") +
  ylab("Greenhous Gas equivalents in tons") +
  ggtitle ("Emission of Greenhous Gas equivalents in tons by account in the European Union in the year 2017")

plot.2017.total.account.barplot

```

Since all the variables already have a long title, the legend is omitted. And because of the long
variable names a flipped bar works better.

The barplot already is really suitable to answer our question. One can clearly spot the main drivers.


#-----------------------------

#Question: How has the account of greenhouse gases in the European Union has changed from 2008 to 2017?

Make a time series of the 4 major following accounts in one graphic
 1. electricity, gas, steam and air conditioning supply
 2. manufacturing
 3. Transportation and storage
 4. Agriculture, forestry and fishing

If we would use more variables the plot might look too confusing with too many graphs.

Subset the data first and change from long to wide format, so every account is in its own column:

```{r}
env_ac_ainah_r2.total.timeseries <- subset(env_ac_ainah_r2, (GEO == "European Union - 28 countries") & ((NACE_R2 == "Electricity, gas, steam and air conditioning supply") | (NACE_R2 == "Manufacturing") | (NACE_R2 == "Transportation and storage") | (NACE_R2 == "Agriculture, forestry and fishing")))                                   

env_ac_ainah_r2.total.timeseries.wide <- dcast(env_ac_ainah_r2.total.timeseries, TIME~NACE_R2, value.var="Value")
```

Now plot the time series. Here time is left as a standard numeric.

```{r}
ggplot() +
  geom_line(data = env_ac_ainah_r2.total.timeseries.wide, aes(x = env_ac_ainah_r2.total.timeseries.wide [,1], y = env_ac_ainah_r2.total.timeseries.wide[,2]),color = "green", size = 2)+
  geom_line(data = env_ac_ainah_r2.total.timeseries.wide, aes(x = env_ac_ainah_r2.total.timeseries.wide [,1], y = env_ac_ainah_r2.total.timeseries.wide[,3]),color = "red", size = 2)+
  geom_line(data = env_ac_ainah_r2.total.timeseries.wide, aes(x = env_ac_ainah_r2.total.timeseries.wide [,1], y = env_ac_ainah_r2.total.timeseries.wide[,4]),color = "blue", size = 2)+
  geom_line(data = env_ac_ainah_r2.total.timeseries.wide, aes(x = env_ac_ainah_r2.total.timeseries.wide [,1], y = env_ac_ainah_r2.total.timeseries.wide[,5]),color = "orange", size = 2)+
  xlab("Time in Years") +
  ylab("Emission in tons") +
  ggtitle ("Time Series of Emission of Greenhous Gas equivalents in tons of different accounts") +
  labs(color = c("A", "B", "C", "D"))
  
```
  
Form: Labels of the axis seem to be missing. The top graph is energy supply, the lower one manufacture.
And those below are Agriculture and Transport.

One can see, that the emission for energy supply and manufacture is sinking, while this is not true
for agriculture and transport.
  

# Question: How similar are different European countries in regard to their account of Greenhouse gas?


First subset again

```{r}
env_ac_ainah_r2.2017 <- subset(env_ac_ainah_r2, (TIME == 2017) & !(GEO == "European Union - 28 countries") & !(NACE_R2 == "Total - all NACE activities"))
```

Problem: we have to reduce the dimensionality. We can just plot two accounts.

First try: Simply use a scatterplot of the biggest two drivers in the European Union in 2017:
 - electricity, gas, steam and air conditioning supply
 - manufacturing

Following the idea, subset even further and change from long to wide again.
Also there are countries, which have no values. These have to be removed.

```{r}
#subset
env_ac_ainah_r2.2majoracc.2017 <- subset(env_ac_ainah_r2.2017, (NACE_R2 == "Electricity, gas, steam and air conditioning supply") | (NACE_R2 == "Manufacturing"))

#from long to wide
env_ac_ainah_r2.2majoracc.2017 <- dcast(env_ac_ainah_r2.2majoracc.2017, GEO~NACE_R2, value.var="Value")

#Remove the countries, where there was NA, but now a 0 was imputed.
#These countries are Iceland, Norway, Serbia, Switzerland and Turkey.

env_ac_ainah_r2.2majoracc.2017 <- subset(env_ac_ainah_r2.2majoracc.2017, !(env_ac_ainah_r2.2majoracc.2017[2] == 0))

str(env_ac_ainah_r2.2majoracc.2017)
```


#Make a scatterplot
```{r}
plot.2majoracc.2017.scatter <- ggplot(env_ac_ainah_r2.2majoracc.2017, aes(x = env_ac_ainah_r2.2majoracc.2017[,2] , y =env_ac_ainah_r2.2majoracc.2017[,3])) +
  geom_point(size=1, shape=23) +
  geom_text(label=env_ac_ainah_r2.2majoracc.2017$GEO) +
  xlab("Emission caused by Energy Supply in tons") +
  ylab("Emission caused by Manufacturing in tons") +
  ggtitle ("Emission of Greenhous Gas equivalents in tons by the 2 major accounts of different European Countries the year 2017")

plot.2majoracc.2017.scatter
```

The names of the countries are barely visible.One can use ggrepel to shift the names to
enhance visibility of the names.
(!ggrepel might move the data points to ensure fewer overlap of the labels)


```{r}
plot.2majoracc.2017.scatter.repel <- ggplot(env_ac_ainah_r2.2majoracc.2017, aes(x = env_ac_ainah_r2.2majoracc.2017[,2] , y = env_ac_ainah_r2.2majoracc.2017[,3])) +
  geom_point(size=1, shape=23) +
  geom_text_repel(label=env_ac_ainah_r2.2majoracc.2017$GEO) +
  xlab("Emission caused by Energy Supply in tons") +
  ylab("Emission caused by Manufacturing in tons") +
  ggtitle ("Emission of Greenhous Gas equivalents in tons by the 2 major accounts of different European Countries the year 2017")

plot.2majoracc.2017.scatter.repel
````


Still it is hard to spot the points for the countries, because most are on one spot.

All in all it is a better idea to scale the greenhouse emissions of the countries for energy supply and
manufacturing by country size. As a quantization population or GDP could be used. Here a scaling is done
via population.

This does not only enhance visibility, but also eases the interpretation. The emission is not only
driven by the specifics of a country, but also by its size.

Load the population countries

```{r}
demo_gind <- read.csv("./Data/Population change - Demographic balance and crude rates at national level (demo_gind)/demo_gind_1_Data.csv")

demo_gind$Value <- gsub(",","",demo_gind$Value) 

demo_gind$Value <- as.double(demo_gind$Value) #NAs introduced by coercion!


demo_gind.2017 <- subset(demo_gind, (TIME == 2017))

#Rename one of the countries
levels(demo_gind.2017$GEO) [levels(demo_gind.2017$GEO) == "Germany (until 1990 former territory of the FRG)"] <- "Germany"
```

Now we have the problem, that we have the population data in one dataset and the emission data in another dataset.
The easiest way to concetenate this data frames is by the inner join of the dplyr package.

```{r}
env_ac_ainah_r2.2majoracc.2017.pop <- inner_join(env_ac_ainah_r2.2majoracc.2017, demo_gind.2017)
```

Inner join seem to have worked. All values of population seem to be reasonable. 
The resulting data frame is not tidy and more important 5 values are missing.
This is neglected for now.

#Add two new columns with energy ghg / population and manufacturing / population

```{r}
env_ac_ainah_r2.2majoracc.2017.pop$GHG_by_energy_supply_per_capita <- env_ac_ainah_r2.2majoracc.2017.pop [,2] / env_ac_ainah_r2.2majoracc.2017.pop [,6]

env_ac_ainah_r2.2majoracc.2017.pop$GHG_by_manufacturing_per_capita <- env_ac_ainah_r2.2majoracc.2017.pop [,3] / env_ac_ainah_r2.2majoracc.2017.pop [,6]
```

#Determine the slope between energy supply and manufacture

In the following scatterplot the countries are plotted with the axis energy supply and manufacture. But
the identity is no good identicator, if a country has more emission from manufacture or from
energy supply, because there is generally more emission from energy supply. This should be taken into
account with an additional line.

```{r}
slope = env_ac_ainah_r2.total.2017$Value [env_ac_ainah_r2.total.2017$NACE_R2 == "Electricity, gas, steam and air conditioning supply"] /env_ac_ainah_r2.total.2017$Value [env_ac_ainah_r2.total.2017$NACE_R2 == "Manufacturing"]
df <- data.frame(x1 = 0, x2 = 2.6, y1 = 0, y2 = 2.6*slope)
```


#Scatterplot of energy supply and manufacture

```{r}
plot.2majoracc.2017.scatter.pop <- ggplot(env_ac_ainah_r2.2majoracc.2017, aes(x = env_ac_ainah_r2.2majoracc.2017.pop[,8] , y = env_ac_ainah_r2.2majoracc.2017.pop[,9])) +
  geom_point(size=2, shape=23) +
  geom_text_repel(label=env_ac_ainah_r2.2majoracc.2017.pop$GEO) +
  geom_segment(aes(x = x1, y = y1, xend = x2, yend = y2, colour = "segment"), data = df) +  
  xlab("Emission caused by Energy Supply in tons per capita") +
  ylab("Emission caused by Manufacturing in tons per capita") +
  ggtitle ("Emission of Greenhous Gas equivalents in tons per capita by the 2 major accounts of different European Countries the year 2017")

plot.2majoracc.2017.scatter.pop
```
One can see that France for example rather tends to have emission from manufacture than from energy supply.
Since the energy supply of France is mostly from nuclear power, this seems to be plausible.

Keep in mind that there is no causality in the graph. As stated in the example with France, the causality is hidden.


#Camparison with shiny
But this are just 2 accounts and there are other importants ones, e.g. agriculture.
One could use the shiny package to compare countries with another and show the barplot
of the accounts from above. (This is omitted here)

#Comparison by dimensionality reduction with principal component analysis

Also you could try to reduce the dimensionality by principal component analysis.Then similar
countries could be clustered on the principal component 1 and 2 plot


Take the data from 2017 for all countries and accounts and change data from long to wide:

```{r}
env_ac_ainah_r2.2017.wide <- dcast(env_ac_ainah_r2.2017, GEO~NACE_R2, value.var="Value")
```

Make a principal component analysis and have a look.

```{r}
env_ac_ainah_r2.2017.wide.pca <- prcomp(env_ac_ainah_r2.2017.wide[,c(2:22)], center = TRUE,scale. = FALSE)

summary(env_ac_ainah_r2.2017.wide.pca)

data.frame(env_ac_ainah_r2.2017.wide.pca$x)
```

Make a skree plot and look how many principal components are important.

```{r}
fviz_eig(env_ac_ainah_r2.2017.wide.pca)
```

So the different accounts of emission can be summarized quite well in one principal component.
This is no surprise, since it's rational to assume that  the emission values from different sources 
are all positively correlated.
On the other hand this means, that a 2 dimensional plot of the countries on the first two principal 
components might not be meaningful. Maybe it's more sensible to plot the countries on just on dimension.

#Nevertheless visualize the first two principle components, where the data points are depicted.
```{r}
fviz_pca_ind(env_ac_ainah_r2.2017.wide.pca, repel = TRUE)
```

Labels are missing, so countries are not easily identifiable.It seems to be the case that a relabeling is not possible, so check the most special ones by hand:

10: France
11: Germany
23: Poland
33: UK

The outliers are mostly big countries. 
Scaling by population was omitted here, but it should not have greater effects.
Other reasons, why these countries are outliers?

This plot was already the goal of the idea to use pca, but in the end the power
of explanation has not seemed to be high.


#How similar are the accounts?
```{r}
fviz_pca_var(env_ac_ainah_r2.2017.wide.pca)
```

Except for acitivities of extraterritorial they mostly point in the same direction.


-----------------------------


Discussion
1. How were the shown plots? Do they answer the question, for which they were made for? 
   Were they good style?
   Have the plots delivered a lot of insight? Which plots haven't?
   What would you have plotted in the field?



